---
title: "Firearms Discussion Classifier"
author: "Bob Horton"
date: "11/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we use the 20-Newsgroups dataset to build a multiclassifier classifier, then use it to detect posts related to firearms. Specifically, we want to use the classifier scores to distinguish posts in `talk.politics.guns` versus posts from all the other newsgroups. For a more difficult problem, we will try to distinguish posts in `talk.politics.guns` from posts in `talk.politics.misc`.


```{r load_libraries, message=FALSE}
library(dplyr)
library(MicrosoftML)
library(ROCR)
library(pROC)
library(ggplot2)
```

```{r load_data}
DATA_DIR <- "text_classification_data"
DATA_FILES <- list(train="train_clean.tsv", test="test_clean.tsv")

data_sets <- DATA_FILES %>% lapply(function(fname){
  read.delim(file.path(DATA_DIR, fname), stringsAsFactors=FALSE, quote='')
})

newsgroups <- unique(data_sets$train$label)

data_sets <- data_sets %>% lapply(function(df){
  df$label = factor(df$label, levels=newsgroups)
  df
})

```

```{r all_vs_all_classifier}

xforms <- list(
  featurizeText(
    vars = c(msg_features="msg"),
    wordFeatureExtractor=ngramCount(ngramLength=1, maxNumTerms=5e5),
    # wordFeatureExtractor=ngramHash(ngramLength=4, hashBits=21),
    stopwordsRemover=stopwordsDefault(), keepPunctuations=FALSE
  )
)

ALPHA <- 0.01
LAMBDA <- 0.5 
L1_WEIGHT <- ALPHA * LAMBDA
L2_WEIGHT <- (1 - ALPHA) * LAMBDA

aa_logreg <- rxLogisticRegression(
    formula = label ~ msg_features,
    type = "multiClass", data = data_sets$train,
    l1Weight=L1_WEIGHT, l2Weight=L2_WEIGHT,
    mlTransforms=xforms)

aa_coef <- coef(aa_logreg)

aa_scores <- rxPredict(aa_logreg, data_sets$test, extraVarsToWrite='label')

plot_roc_for_newsgroup <- function(scores_df, newsgroup_name){
  score_col <- sprintf('Score.%s', newsgroup_name)
  score <- scores_df[[score_col]]
  label <- scores_df$label == newsgroup_name
  pred_obj <- prediction(score, label)
  roc_obj <- performance(pred_obj, 'tpr', 'fpr')
  auc <- performance(pred_obj, 'auc')@y.values[[1]]
  plot(roc_obj, main=newsgroup_name)
  abline(0, 1, lty=3, col='blue')
  text(0.6, 0.2, sprintf('AUC: %0.3f', auc))
}

```

```{r plot_roc_curves}

for (newsgroup in newsgroups){
  plot_roc_for_newsgroup(aa_scores, newsgroup)
  aa_coef[grepl(newsgroup, names(aa_coef))] %>% head(20) %>% print
}

```

## Confusion Matrix

```{r confusion_matrix}
M <- with(aa_scores, table(label, PredictedLabel)) %>% as.matrix

sum(M * diag(nrow(M)))/sum(M) # accuracy

column_totals <- aa_scores %>% 
  group_by(label) %>% 
  summarize(col_total=n())

aa_scores %>% 
  select(label, PredictedLabel) %>% 
  group_by(label, PredictedLabel) %>% 
  summarize(frequency=n()) %>% 
  left_join(column_totals, by='label') %>% 
  mutate(relative_frequency = frequency/col_total) %>% 
  ggplot(aes(x=label, y=PredictedLabel, fill=relative_frequency)) + 
    geom_tile() +
    scale_x_discrete(name="Actual Class") + 
    scale_y_discrete(name="Predicted Class") + 
    theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5)) +
  scale_fill_gradient(low="white", high="darkorange") + labs(fill="Relative Frequency")

```

## Guns vs. Politics


How well can we distinguish posts in `talk.politics.guns` from posts in `talk.politics.misc`?

```{r guns_vs_misc}
keepers <- c('talk.politics.guns', 'talk.politics.misc')

guns_and_misc <- aa_scores %>% 
  filter(label %in% keepers) %>% 
  select(label, Score.talk.politics.guns, Score.talk.politics.misc) %>% 
  mutate(label=factor(as.character(label))) %>% 
  arrange(desc(Score.talk.politics.guns/Score.talk.politics.misc))

roc1 <- with(guns_and_misc, roc(label, Score.talk.politics.guns))
auc(roc1)

roc2 <- with(guns_and_misc, roc(label, (Score.talk.politics.guns/Score.talk.politics.misc)))
auc(roc2)

# add a very small bit of noise to avoid ties
roc3 <- with(guns_and_misc, roc(label, (Score.talk.politics.guns/Score.talk.politics.misc)+rnorm(nrow(guns_and_misc), sd=0.0001)))
auc(roc3)



pred_obj <- with(guns_and_misc, ROCR::prediction(
  (Score.talk.politics.guns/Score.talk.politics.misc)+rnorm(nrow(guns_and_misc), sd=0.0001),
  label=='talk.politics.guns'
))

perf_roc <- ROCR::performance(pred_obj, 'tpr', 'fpr')

ss3 <- data.frame(TPR=perf_roc@y.values[[1]][-1],
                  FPR=perf_roc@x.values[[1]][-1],
                  label=guns_and_misc$label=='talk.politics.guns')

ss3 %>% ggplot(aes(x=FPR, y=TPR, col=label)) + geom_point(alpha=0.5)

plot(perf_roc)
with(ss3, points(FPR, TPR, col=ifelse(label, 'red', 'black'), pch=20))
abline(0, 1, lty=3, col='blue')


```

```{r save_guns_roc, eval=FALSE, echo=FALSE}
guns_roc <- with(aa_scores, roc(label=='talk.politics.guns', Score.talk.politics.guns))
plot(guns_roc, print.auc=TRUE)
saveRDS(guns_roc, 'guns_roc.Rds')
```